---
title:  "SYS 5581 Time Series and Forecasting"
author: "Computing setup and guide"
date:   "University of Virginia Engineering, Spring 2021"

output: 
  pdf_document:
#   keep_tex: true
    fig_caption: yes
    latex_engine: pdflatex

  # html_document:
  #   keep_md: true

urlcolor: blue
geometry: margin=1in

fontfamily: mathpazo
fontsize: 11pt
header-includes:
   - \linespread{1.05}
---

```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.path='figs/')
knitr::opts_chunk$set(cache.path='_cache/')
knitr::opts_chunk$set(warning=F, message=F)
# knitr::opts_knit$get(kable.force.latex = TRUE)
```



```{r dates_bib, include=FALSE}
# knitr::opts_chunk$set(cache=FALSE, dev='pdf')
# mon <- as.Date("2016-08-15")
# # fri <- as.Date("2016-01-08")
advdate <- function(obj, adv) {}
# advdate <- function(obj, adv) {
#  tmon <- obj + 7*(adv-1)
#  tfri <- obj + 4 + 7*(adv-1)
#  tmon <- format(tmon, format="%m/%d")
#  tfri <- format(tfri, format="%m/%d")
#  zadv <- sprintf("%02d", adv)
#  tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri)
#  return(tmp)
# }
# library(RefManageR)
# # library(knitcitations)
# # library(rcrossref)
# bib <- ReadBib("master.bib")
# myopts <- BibOptions(bib.style = "authoryear", style="latex", first.inits=FALSE, max.names = 20)
```

The course relies on computing resources. Please install the software as indicated on your local machine, and familiarize yourself with the associated documentation. 

# The R programming language

* [Download and install R](https://cran.rstudio.com/), v. 3.0.1+.
* [Download and install R Studio](https://rstudio.com/products/rstudio/download/#download), v. 1.4.1+. 
  - R Studio is an integrated development environment (IDE) for R. It offers a variety of utilities to enhance the experience of coding and generating documents.
* Install the [Tidyverse packages for R](https://www.tidyverse.org/): From the Console tab in R Studio (or from R running in a Terminal window), enter: `install.packages("tidyverse")`.
  - Tidyverse is a collection of packages that extend the capabilities of R for doing data science.
* Install the [tidyverts packages for R](https://tidyverts.org/): From the Console tab in R Studio (or from R running in a Terminal window), enter: `install.packages("tidyverts")`.
  - tidyverts is a collection of R packages for time series analysis designed to work well with the Tidyverse packages.

# Git and Github

[Git](https://git-scm.com/) is software for version control. Github is a web service that provides remote storage and access to files via git. This setup greatly facilitates collaboration between multiple individuals working on the same code base.

* [Follow these instructions](https://jennybc.github.io/2014-05-12-ubc/ubc-r/session03_git.html) to [download and install git](https://git-scm.com/downloads) and to link git with R Studio.

A collection of files associated with a single project is in git-speak called a "repository" or *"repo"*. You should already have a basic repo set up for you on the course site on Github. The next step is to copy ("clone") this remote repo to your local machine.

* Clone your course repo on Github to a new R Studio project on your local machine.
  - Navigate to [the course website on Github](https://github.com/uva-eng-time-series-sp21). Select your repo.
  - Click on the green button labeled "Code". Copy the URL.
  - In the R Studio window, from the pull-down menu in the upper-right corner, select `New Project...`, `Version Control`, `Git`. Paste the URL into the dialog box labeled `Repository URL`.
  - Optional: Change the name of the project folder, and the location of this folder on your local directory tree.
  - Click on `Create Project`. The files from your remote repo should be copied to your local machine in a new folder with the name you chose.
  
* Optional: [Download and install the Github desktop client](https://git-scm.com/downloads/guis), or an alternative GUI client.
  - The git operations you need for this course can be managed within R Studio, from the `Git` tab. Some more advanced operations require using either a Terminal window, or a Git desktop client. 
  
* [Review the documentation for git](https://git-scm.com/) and [Github](https://guides.github.com/). Learn the basics.


# Markdown and R Markdown



[Coming soon...]

https://www.markdownguide.org/

https://rmarkdown.rstudio.com/

# Bibliographic resources: Zotero and Bibtex

[Coming soon...]

# General course web resources

  * [Collab class site](https://collab.its.virginia.edu), for basic course information, assignments, office hours sign-up, links to online textbook and other resources.
  * [Github class site](https://github.com/uva-eng-time-series-sp21), for posting and sharing code.
  * Zoom, for class sessions, recordings, and office hours.



