---
title:    "Outline of topics"
subtitle: "SYS 5581 Time Series and Forecasting"
author: "Instructor: Arthur Small"
date:   "University of Virginia Engineering, Spring 2021"

output: 
  pdf_document:
#   keep_tex: true
    fig_caption: yes
    latex_engine: pdflatex

  # html_document:
  #   keep_md: true

urlcolor: blue
geometry: margin=1in

fontfamily: mathpazo
fontsize: 11pt
header-includes:
   - \linespread{1.05}

bibliography: /Users/Arthur/GitRepos/Teaching/time-series/tseries.bib
---

```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.path='figs/')
knitr::opts_chunk$set(cache.path='_cache/')
knitr::opts_chunk$set(warning=F, message=F)
# knitr::opts_knit$get(kable.force.latex = TRUE)
```

```{r dates_bib, include=FALSE}
# knitr::opts_chunk$set(cache=FALSE, dev='pdf')
# mon <- as.Date("2016-08-15")
# # fri <- as.Date("2016-01-08")
# advdate <- function(obj, adv) {}
# advdate <- function(obj, adv) {
#  tmon <- obj + 7*(adv-1)
#  tfri <- obj + 4 + 7*(adv-1)
#  tmon <- format(tmon, format="%m/%d")
#  tfri <- format(tfri, format="%m/%d")
#  zadv <- sprintf("%02d", adv)
#  tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri)
#  return(tmp)
# }
# library(RefManageR)
# # library(knitcitations)
# # library(rcrossref)
# bib <- ReadBib("master.bib")
# myopts <- BibOptions(bib.style = "authoryear", style="latex", first.inits=FALSE, max.names = 20)
```

FPP = [Hyndman, R.J., & Athanasopoulos, G. [-@robjhyndmanForecastingPrinciplesPractice2021] *Forecasting: principles and practice, 3rd edition*, OTexts: Melbourne, Australia. OTexts.com/fpp3](https://otexts.com/fpp3/)

# Introduction and setup

Topics:

* What is time series analysis?


Readings: 

* FPP, Ch. 1

Assignment: Write a concept note for a potential project.

## Computing setup

Topics: R, R Studio, git, Github, Markdown, R Markdown.

Assignment: Follow instructions in the course [Computing setup guide](https://github.com/uva-eng-time-series-sp21/sys5581-course-materials/blob/master/computing_setup_guide.pdf).


# Data extraction, transformation, and loading

Topics:

* Tidy data
* Tidyverse and tidyverts packages

Readings:

* FPP, Section 2.1

Assignment: [Extract, transform, and load your data](https://github.com/uva-eng-time-series-sp21/sys5581-course-materials/blob/master/assignments/assignment_data_etl.pdf).

# Exploratory data analysis

Readings: FPP, Ch. 2 and Ch. 4

Assignment: Explore your data





# Planning a time series analysis project




## References