---
title: 'Class notes: Model Selection, continued'
author: 'Instructor: Arthur Small'
date: '2021-03-24'
output:  
  slidy_presentation: default
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 4
    code_folding: show
    fig_caption: yes
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
institute: SYS 5581 Time Series & Forecasting, Spring 2021
bibliography: /Users/Arthur/GitRepos/Teaching/time-series/tseries.bib
link-citations: yes
---

# Reminders, comments

* NO CLASS MONDAY, MARCH 29
  - Enjoy your break

* Reminder: revised deadlines:
  - Weds Mar 24: Project proposal
     - Question
     - Data
     - Data prep
     - Data exploration, extraction of salient features
  - Weds Mar 31: Model
  
* Comment on Github: how to process pull requests
  - Esp. those with instructor's with comments & suggestions

# Last time: Model selection

* Linear regression models
  - Transformations
  
* Introduce ARIMA(p,d,q) models

# Today:

* Working with modeling functions in `fable` package: 
  - TSLM() ("time series linear model") for linear regressions on time series data
  - Reference: FPP3, Chapter 7


* [Time permitting:] Adding bibliographies and citations to Markdown documents
  - Zotero for managing libraries of resources: web, desktop, browser plugin
  - Bibtex for generating bibliographies in documents: .bib files, BibTeX, bibdesk


<span style="color: blue">Part I</span>
===

<span style="color: blue">*Using fable::TSLM() to estimate time series linear models*</span>


```{r include=FALSE}
library(tidyverse)
library(fpp3)
library(tsibbledata)
```


## 

```{r}
# print(us_change)
# 
# 
# ggplot(data = us_change, aes(Quarter, Consumption,)) +
#   geom_line() 
# 
# 
# +
#   geom_line(us_change, aes(y =Income))
```


<!-- Why fit statistical (regression) models? -->
<!-- === -->

<!-- You have some data $X_1,\ldots,X_p,Y$: the variables $X_1,\ldots,X_p$ are called predictors, and $Y$ is called a response. You're interested in the relationship that governs them -->

<!-- So you posit that $Y|X_1,\ldots,X_p \sim P_\theta$, where $\theta$ represents some unknown parameters. This is called **regression model** for $Y$ given $X_1,\ldots,X_p$. Goal is to estimate parameters. Why? -->

<!-- - To assess model validity, predictor importance (**inference**) -->
<!-- - To predict future $Y$'s from future $X_1,\ldots,X_p$'s (**prediction**) -->



<!-- # Model selection: Regression models: Summary -->

<!-- You want to explain observed $y_1, \ldots, y_T$, and predict future $y_{T+1}, y_{T+2}, \ldots$. -->

<!-- Observed values may depend on past values of $y$, and/or on other predictors $X_1,\ldots,X_K$. -->





<!-- #  -->



