---
title:     "Exploratory analysis of time series data: Examples"
institute: "SYS 5581 Time Series & Forecasting, Spring 2021" 
author:     "Instructor: Arthur Small"
date:       "Version of `r Sys.Date()`"
output: 
  html_notebook:
    number_sections: true
    toc: yes
    toc_depth: 4
    code_folding: show # options: show, hide
    fig_caption: yes
  # html_document:
  #       keep_md: yes
  # pdf_document: default
bibliography: /Users/Arthur/GitRepos/Teaching/time-series/tseries.bib
link-citations: yes
---

```{r set up coding environment, include=FALSE, message=FALSE}
# library(dplyr) -- don't need this if you are loading the entire 'tidyverse' suite
library(tidyverse)
library(lubridate) # For easy handling of time-indexed objects

# if(!('fpp3' %in% installed.packages())) install.packages('fpp3')
library(fpp3)
```


# Example: Monthly electricity sales for Virginia

## Extract data from remote database

```{r open connection to database, eval=FALSE, include=FALSE}
# Open connection to a remote database
# Make sure your VPN network connection is active if needed!

# if(!('RPostgreSQL' %in% installed.packages())) install.packages('RPostgreSQL')
library(RPostgreSQL)

# "my_postgres_credentials.R" contains the log-in information
source("/Users/Arthur/GitRepos/Teaching/my_postgres_db_credentials.R")

# Open connection
db_driver <- dbDriver("PostgreSQL")
db <- dbConnect(db_driver,user=user, password=password,dbname="postgres", host=host)
rm(password) 

# check the connection: If function returns value TRUE, the connection is working
dbExistsTable(db, "metadata")
```

```{r retrieve data from db, eval=FALSE, message=FALSE}
esales <- dbGetQuery(db,'SELECT * from eia_elec_sales_va_all_m') # SQL code to retrieve data from a table in the remote database
# str(esales)
esales <- as_tibble(esales) # Convert dataframe to a 'tibble' for tidyverse work
# str(esales)
```

```{r save data in Apache Arrow format, eval=FALSE}
# Reference: https://arrow.apache.org/docs/r/
# if(!('arrow' %in% installed.packages())) install.packages('arrow')
library(arrow)
write_feather(esales, "esales.feather")
```

```{r close db connection, eval=FALSE}
# Close connection -- this is good practice
dbDisconnect(db)
dbUnloadDriver(db_driver)
```

## Briefly characterize the dataset

```{r read in data}
library(arrow)
esales <- read_feather("esales.feather")
```

```{r }
print(esales)    # print the data as a table
summary(esales)  # compute basic summary statistics about the data
boxplot(esales)
```

```{r use tidyverse syntax to perform some simple data manipulations}
# References: https://www.tidyverse.org/, https://dplyr.tidyverse.org/
# filter(data object, condition) : syntax for filter() command

esales %>%
  filter(year == 2019) %>%
  filter(value > 9000) %>%
  print()

(esales %>%
  group_by(year) %>%
  summarise(Total = sum(value)) -> total_esales_by_year)

esales %>%
  mutate(sales_TWh = value/1000) %>%
  select(-value)
```

## Plot the time series

```{r use ggplot2 to generate a plot}
#Reference: https://ggplot2.tidyverse.org/

ggplot(data=esales, aes(x=date,y=value)) +
  geom_line() + xlab("Year") + ylab("Virginia monthly total electricity sales (GWh)")
```

```{r}
# library(lubridate) # Make it easy to deal with dates

esales %>% filter(month==3)                   # These three lines of code
esales %>% filter(month(date)==3)             #   all do
esales %>% filter(lubridate::month(date)==3)  #   the same thing.

# We don't have to keep the 'year' and 'month' column: can recover them if needed

esales %>%
  select(date, sales_GWh = value) -> esales_tbl

print(esales_tbl)
```

## Convert the data frame into a time series `tsibble` object


```{r}
# install.packages("tsibble")
library(tsibble) # Reference: https://tsibble.tidyverts.org/articles/intro-tsibble.html

esales_tbl %>% as_tsibble(index = date) -> elsales_tbl_ts

print(elsales_tbl_ts)
```

## Make some plots

###  Make a histogram of monthly sales

```{r make a histogram of the data}
hist(elsales_tbl_ts$sales_GWh, breaks=40)
```

###  Make a seasonal plot

```{r, eval=FALSE}
# This plot won't work. Why not?
elsales_tbl_ts %>% 
  feasts::gg_season(sales_GWh, labels = "both") + ylab("Virginia electricity sales (GWh)")
```

```{r Change the tsibble so index is monthly}
# install.packages("feasts"), Reference: https://feasts.tidyverts.org/
library(feasts)

elsales_tbl_ts %>% 
  mutate(Month = tsibble::yearmonth(date)) %>% 
  as_tsibble(index = Month) %>% 
  select(Month,sales_GWh) -> vaelsales_tbl_ts

print(vaelsales_tbl_ts)
```
```{r}
autoplot(vaelsales_tbl_ts, sales_GWh) + 
  ylab("Virginia monthly total electricity sales (GWh)") + 
  xlab("")  # Leave horiz. axis label blank
```
## Seasonal plots and seasonal subseries plots

```{r , warning=FALSE}
vaelsales_tbl_ts %>% gg_season(sales_GWh, labels = "both") + ylab("Virginia electricity sales (GWh)")
```


```{r}
vaelsales_tbl_ts %>% 
  gg_subseries(sales_GWh) +
  labs(
    y = "Sales (GWh)",
    title = "Seasonal subseries plot: Virginia electricity sales"
  )
```

## Scatterplots

Investigating relationships between two variables. Scatterplots. Correlation. Scatterplot matrices.

Readings: FPP Sect. 2.6

```{r}
vic_elec

summary(vic_elec)

vic_elec %>%
  filter(year(Time) == 2013) %>%
  autoplot(Demand) +
  labs(
    y = "Demand (GW)",
    title = "Half-hourly electricity demand: Victoria"
  )
```
```{r}
vic_elec %>%
  filter(year(Time) == 2013) %>%
  autoplot(Temperature) +
  labs(
    y = "Temperature (degrees Celsius)",
    title = "Half-hourly temperatures: Melbourne, Australia"
  )
```

```{r}
vic_elec %>%
  filter(year(Time) == 2013) %>%
  ggplot(aes(x = Temperature, y = Demand)) + 
#  geom_density2d() +
  geom_point(size=0.1, aes(colour=Holiday), alpha = 0.4) +
  labs(y = "Demand (GW)", x = "Temperature (degrees Celsius)")
```
A Scatterplot matrix

```{r, warning=FALSE}
vic_elec

boxplot(vic_elec$Temperature)

# install.packages("GGally")
vic_elec %>% 
  # mutate(Temperature = round(Temperature)) %>%
  # pivot_wider(values_from=c(Demand,Temperature), names_from=Holiday) %>%
  GGally::ggpairs(columns = 3:2)

vic_elec %>% 
  mutate(Year = factor(year(Date))) %>%
  select(-c(Date, Holiday)) %>%
  GGally::ggpairs(columns = 4:2)
```

# Example: Australian production

```{r, warning=FALSE}
# install.packages('tsibbledata')
library(tsibbledata)

aus_production

aus_production %>% gg_season(Electricity)

aus_production %>% gg_season(Beer)
```

# Example: Gross Domestic Product data

## Exploratory data analysis

```{r}
library(tsibbledata) # Data sets package

print(global_economy)
```

```{r}
global_economy %>% filter(Country=="Sweden") %>% print()
```

```{r}
global_economy %>%
  filter(Country=="Sweden") %>%
  autoplot(GDP) +
  ggtitle("GDP for Sweden") + ylab("$US billions")
```

## Fitting data to simple models

```{r}
global_economy %>% model(trend_model = TSLM(GDP ~ trend())) -> fit

fit


```

```{r}
fit %>% filter(Country == "Sweden") %>% residuals()
```

```{r}

fit %>% filter(Country == "Sweden") %>% residuals() %>% autoplot(.resid)
```

### Work with ln(GDP)

```{r}
global_economy %>%
  filter(Country=="Sweden") %>%
  autoplot(log(GDP)) +
  ggtitle("ln(GDP) for Sweden") + ylab("$US billions")
```

```{r}
global_economy %>%
  model(trend_model = TSLM(log(GDP) ~ trend())) -> logfit
```

```{r}
logfit %>% filter(Country == "Sweden") %>% residuals() %>% autoplot()
```

```{r}
global_economy %>% model(trend_model = TSLM(log(GDP) ~ log(Population))) -> fit3

fit3 %>% filter(Country == "Sweden") %>% residuals() %>% autoplot()

```

# Producing forecasts

```{r}
fit %>% forecast(h = "3 years") -> fcast3yrs

fcast3yrs

```

```{r}

fcast3yrs %>% filter(Country == "Sweden", Year == 2020) %>% str()
```

```{r visualize forecasts}
fcast3yrs %>% 
  filter(Country=="Sweden") %>%
  autoplot(global_economy) +
  ggtitle("GDP for Sweden") + ylab("$US billions")
```

## Model residuals vs. forecast errors

Model residuals:

Your data: $y_1, y_2, \ldots, y_T$

Fitted values: $\hat{y}_1, \hat{y}_2, \ldots, \hat{y}_T$

Model residuals: $e_t = y_t - \hat{y}_t$

Forecast errors:

```{r}
augment(fit)
```

```{r}
augment(fit) %>% filter(Country == "Sweden") %>%
  ggplot(aes(x = .resid)) +
  geom_histogram(bins = 20) +
  ggtitle("Histogram of residuals")
```

## Are the model residuals auto-correlated?

```{r}
augment(fit) %>% filter(Country == "Sweden") -> augSweden

augSweden %>%
  ACF(.resid) %>%
  autoplot() + ggtitle("ACF of residuals")
```

```{r}
augment(fit3) %>% filter(Country == "Sweden") -> augSweden3

augSweden3 %>%
  ACF(.resid) %>%
  autoplot() + ggtitle("ACF of residuals")
```


# Example: GDP, several countries


```{r}
library(tsibbledata) # Data sets package

nordic <- c("Sweden", "Denmark", "Norway", "Finland")

(global_economy %>% filter(Country %in% nordic) -> nordic_economy)

```

```{r}
nordic_economy %>% autoplot(GDP)
```

```{r}
fitnord <- nordic_economy %>%
  model(
    trend_model = TSLM(GDP ~ trend()),
    trend_model_ln = TSLM(log(GDP) ~ trend()),
    ets = ETS(GDP ~ trend("A")),
    arima = ARIMA(GDP)
  )

fitnord
```

```{r}
fitnord %>%
  select(arima) %>%
  coef()
```


Denmark: ARMA(1,1)

Finland: MA(2)

Norway: MA(1)

Sweden: MA(2)

```{r}
nordic_economy %>%
  model(arima_constrained = ARIMA(GDP ~ pdq(1,0,2))) %>% select(arima_constrained) %>% coef()
```

```{r}
fitnord %>% coef() 
```

```{r}
fitnord %>%  glance()  
```

```{r}
fitnord %>% filter(Country == "Denmark") %>% select(arima) %>% report()
```

```{r}
fitnord %>%
  accuracy() %>%
  arrange(Country, MPE)
```



```{r, eval=FALSE}
# ETS forecasts
USAccDeaths %>%
  ets() %>%
  forecast() %>%
  autoplot()
```

```{r, eval=FALSE}
str(taylor)
plot(taylor)
```

## Plot lagged values

```{r plot lagged values}
vaelsales_tbl_ts  %>% filter(month(Month) %in% c(3,6,9,12)) %>% gg_lag(sales_GWh, lags = 1:2)

vaelsales_tbl_ts  %>% filter(month(Month) == 1) %>% gg_lag(sales_GWh, lags = 1:2)
```

```{r}
vaelsales_tbl_ts %>% ACF(sales_GWh) %>% autoplot()
```

```{r perform automated time series decomposition}


# decompose(vaelsales_tbl_ts)
```

```{r perform additive STL decomposition of the VA electricity sales time series}
vaelsales_tbl_ts %>%
  model(STL(sales_GWh ~ trend(window=21) + season(window='periodic'), robust = TRUE)) %>%
  components() %>%
  autoplot()
```

```{r perform multiplicative STL decomposition of the VA electricity sales time series}
vaelsales_tbl_ts %>%
  mutate(ln_sales_GWh = log(sales_GWh)) %>%
  model(STL(ln_sales_GWh ~ trend(window=21) + season(window='periodic'),
    robust = TRUE)) %>%
  components() %>%
  autoplot()
```

```{r}
vaelsales_tbl_ts %>%
  features(sales_GWh, feat_stl)
```

```{r}
vaelsales_tbl_ts %>%
  features(sales_GWh, feature_set(pkgs="feasts"))
```



# References
